Typed Functional Genetic Programming

### 2013-07-29 tex/thesis.tex
uterý rano

It scored 17/50 (34\%) success rate. 
Minimal $I(M,i,z)$ was in generation 23 
with 192,000 individuals to be processed.
The experiment took 46 minutes.

@@

\includegraphics[scale=0.65]{reports/SSR/1/indivs.png}

@@

[todo OBRAZKY GRAFů]

@@

@@

The Santa Fe trail has following layout.

@@

Action $ifa x y$ performs action $x$ if a food piece is ahead of the ant,
otherwise it performs action $y$.

@@

Tento problém je také převzatý z \cite{koza92}.

Jde v něm o to [podivat ze do kozy a převypravet to]

sigma a gama jsou

fitnes je následující

koza uvádí že jedinec [ten jedinec] je jedinec který mu vznikl
při vyšlechtění. Na našem systému ale trvá mu to nakejch tech 500něco
krků přičemž by to mělo trvat 400 max.

Nevíme zda je chyba na mé straně (nepřesná implementace) nebo
se jedná o nepřesnost v koza92. Díky tomuto rozporu však nepoužijeme
tuto ulohu pro dokladování toho že se to chová jako 
standard GP. My jsme tedy oproti knize použili max počet kroků 600
tak aby popsaný jedinec v koza92 byl korektní i vnašem systému.

Uvádíme pokus o 50 behach s (geom strategie default q, best preservation on)
...

>>>>

It scored 17/50 (34\%) success rate. 
Minimal $I(M,i,z)$ was in generation 23 
with 192,000 individuals to be processed.
The experiment took 46 minutes.\\

\includegraphics[scale=0.65]{reports/SSR/1/indivs.png}\\

@@

@@

\includegraphics[scale=0.65]{reports/SSR/2/probabs.png}

\includegraphics[scale=0.65]{reports/SSR/2/indivs.png}\\

@@

~\\TODO dát sem ty grafy průměrný term size

@@

The Santa Fe trail lies on toroidal square grid and has following layout.

@@

Action $ifa x y$ (if-food-ahead) performs action $x$ if a food piece is ahead of the ant,
otherwise it performs action $y$.

@@

Ants action is performed over and over again until it reaches predefined
maximal number of steps. In \cite{koza92} this limit is set to be 400 time steps.

Fitness value is equal to number of eaten food pieces.

An individual such that eats all 89 pieces of food is 
considered as a correct solution.\\

There is also mentioned following ant program which is described as correct solution.

\texttt{~\\
(ifa m (p3 l (p2 (ifa m r) (p2 r (p2 l r)))(p2 (ifa m l) m)))\\
}

But here arises a problem: In our implementation this program needs 545
time steps; if it is given only 400 time steps, then it eats only 79 pieces
of food. It is uncertain whether it is caused by my mistake in implementation
or by inaccuracy in \cite{koza92}. Since everything else seems working just 
fine, I have decided to set the limit to 600 in my experiments.

Due to those uncertainties, we will not perform comparing experiments 
as we did in the problem of simple symbolic regression.\\

In the following experiments, the option of preserving the best individual 
into the next generation is active (since we are not trying to compare 
with results in \cite{koza92}).

As usually, first experiment consisted of 50 independent runs of GP algorithm.
Each run had maximally 50 generations and 500 individuals as population size.
\textit{Ramped half-and-half strategy} is used as search strategy for the first 
experiment.\\

It scored 7/50 (14\%) success rate. 
Minimal $I(M,i,z)$ was in generation 28 
with 449,500 individuals to be processed.
The experiment took 265 minutes.

Similarly as in previous problem, second experiment was performed to compare 
\textit{ramped half-and-half strategy} with our
\textit{geometric strategy} (with default parameter $q=0.75$). 
Thus all control parameters stay same except for the 
search strategy for generating algorithm.

It scored 19/50 (38\%) success rate. 
Minimal $I(M,i,z)$ was in generation 10 
with 115,500 individuals to be processed.
The experiment took 107 minutes.\\

\includegraphics[scale=0.65]{reports/Ant/1/probabs.png}

\includegraphics[scale=0.65]{reports/Ant/1/indivs.png}\\

This is big improvement in all three watched factors:

\begin{enumerate}
 \item 38\% vs 14\% success rate.
 \item 115,500 vs 449,500 individuals needed to be processed.
 \item 107 vs 265 minutes.  
\end{enumerate}

Third factor, the time difference might give us clue about what is going on.

Those two experiments differs only in used term generating method.
During one run the vast majority of the time is spend by evolution,
the generation process only slightly prolongs the time spend in generation 0.
The run time of the evolution is influenced by two important factors:
Number of generations and time that crossover takes.

The number of generations of more successful method is surly smaller,
but it seems improbable that it would have such a big impact 
(this is idea is supported by results of experiments with
even-parity problem, where times are almost the same but success 
rates differs significantly).

Time of crossover is proportionate to number of nodes that 
an individual has. Thus it seems that observed slowdown is
due to terms being too big. \\

In order to test this hypothesis, we run additional experiments
to get statistics about dynamics of individuals' sizes.

~\\TODO dát sem ty grafy průměrný term size
